---
title: "Project RD"
author: "John Gleason"
date: "5/4/2022"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Intro 
## What: 
I would like to look through and examine PGA Tour stats from 2019 to 2015 to shed light on what area of the game effective the player score and world rankings the most. 

## Why:
A new approach to the game of golf on the PGA tour has caught some momentum in the last few years. Historically, the thought was to score well on tour players must be extremely efficient on the putting green. In today game, there are a hand full of players that are beginning to focus more on driving distance than putting efficiency. I want to dig deeper into which approach is backed by the numbers.

## How:
I am going to pull 5 years of data from pgatour.com site for the top 100 players in the world. I am going to be looking Driving Distance, Green in Regulation and putting strokes per round.

# Body 
# Topics From Class 

## Topic 1:
Installing and Cleaning of Data
```{r}
library(readxl)
library(tidyverse)
library(stringr)
```


```{r}
pgafulldata_1_copy <- read_excel("~/Desktop/pgafulldata (1) copy.xlsx")
```

```{r}
pga_data_2 <- pgafulldata_1_copy %>% na.omit()
```


```{r}
pga_data_2$Year.Date <- str_sub(pga_data_2$'PLAYER NAME', -6)
```


```{r}
pga_data_2$PLAYER.NAME <- str_sub(pga_data_2$'PLAYER NAME',0 ,-7)

```


```{r}
PGA <- pga_data_2[-c(1,3,6,8,9,10,11)]
```




## Topic 2:
Understanding the histograms of the variable and if their is linear regression 
```{r}
summary(PGA)
```

```{r}
ggplot(PGA) + geom_histogram(aes(x=DrivingDistance), binwidth = 1)
```


```{r}
ggplot(PGA) + geom_histogram(aes(x=GIR), binwidth = 1)
```


```{r}
ggplot(PGA) + geom_histogram(aes(x=PuttsPerRound))
```


```{r}
plot(FedexCupPoints ~ PuttsPerRound + GIR + DrivingDistance, data = PGA)
```



## Topic 3: 
Filtering and using Mean to Compare 
```{r}
Q3 <- PGA %>% filter(FedexCupPoints > 833)
```


```{r}
Q1 <- PGA %>% filter(FedexCupPoints < 323)
```


```{r}
summary(Q3)
```


```{r}
summary(Q1)
```


```{r}
DDPC<- ((297.4-289.4)/289.4)*100
```

```{r}
DDPC
```


```{r}
PPRPC <-((28.94 - 29.22)/29.22)*100
```

```{r}
PPRPC
```


```{r}
GIRPC <- ((78.51 - 75.51)/ 75.51)*100
```

```{r}
GIRPC
```


# Conclusion
